Manuscript title here

### 2024-09-16 manuscript.Rmd
bättre ord
tog bort en ställe som krävde källa, för att det inte krävs

Trauma is a major global public health concern. It causes over four milion death and affects hundreds of millions each year, among those younger populations are the most affected [@noauthor_injuries_nodate]. Shock is a major cause of mortality in trauma patients. The definition for shock is circulatory failure leading to insuficient perfusion and oxygenation of tissue and organ[@leech_shock_2023]. There are four types of causes, which are hypovolemic, obstructive, cardiogenic and distributive. These causes can occur alone or in combination. The most usual one during trauma are hypovolemic in form of a large bleeding[@teixeira_preventable_2007].

@@

The initial management of trauma patients once they reach hospital is very time sensitive and error prone [@mackersie_pitfalls_2010]. To improve trauma care quality, healthcare systems identify Opportunities for Improvement (OFI) through structured reviews of patient cases. Multidisciplinary morbidity and mortality reviews play a central role in this process, where clinicians from different specialities together analyse trauma cases to evaluate both the structure and clinical processes involved in trauma care following the Donabedian’s quality of care framework[@albaaj_patient_2023]. For shock patients, timely interventions such as fluid resuscitation/blood transfusion, intervention, OR and radiology are critical, and delays or missteps can result in adverse outcomes (källa). By systematically addressing OFIs, healthcare providers can improve the effectiveness of trauma care, directly impacting mortality and morbidity(källa).

>>>>

Trauma is a major global public health concern. It causes over four milion death and affects hundreds of millions each year, among those younger populations are the most affected [@noauthor_injuries_nodate]. Shock is a major cause of mortality in trauma patients. The definition for shock is circulatory failure leading to insuficient perfusion and oxygenation of tissue and organ[@leech_shock_2023]. There are four types of causes, which are hypovolemic, obstructive, cardiogenic and distributive. These causes can occur alone or in combination. The most usual one during trauma are hypovolemic in form of a major haemorrhage [@teixeira_preventable_2007].

@@

The initial management of trauma patients once they reach hospital is very time sensitive and error prone [@mackersie_pitfalls_2010]. To improve trauma care quality, healthcare systems identify Opportunities for Improvement (OFI) through structured reviews of patient cases. Multidisciplinary morbidity and mortality reviews play a central role in this process, where clinicians from different specialities together analyse trauma cases to evaluate both the structure and clinical processes involved in trauma care following the Donabedian’s quality of care framework[@albaaj_patient_2023]. For shock patients, timely interventions such as fluid resuscitation/blood transfusion, intervention, OR and radiology are critical, and delays or missteps can result in adverse outcomes (källa). By systematically addressing OFIs, healthcare providers can improve the effectiveness of trauma care, directly impacting mortality and morbidity.
